---
title: "Lab 02 - Plastic waste"
author: "Enhui"
date: "01/20/2025
"
output: github_document
---

## Load packages and da
a

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

The distribution is right-skewed, suggesting that most countries produce less plastic waste per capita, and a few countries produce more waste. Also, the disparity in plastic waste is large between countries. 

```{r plastic-waste-continent}

ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)

plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)

```

### Exercise 2

Because color and fill represent different variables, we want the color to represent different continents in the dataset. Alpha is used to represent the transparency of the data and does not indicate any variable differences in the data. 

```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.7)
```

### Exercise 3

Violin plots show the peaks, which indicate where the data is concentrated and the distribution of the data, but do not show in box plots. Box plots show the median, quartiles, and outliers explicitly, but violin plots do not explicitly show these statistics. 

```{r plastic-waste-violin}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_boxplot()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_violin()
```

### Exercise 4

4.1 Based on the below scatterplot, I cannot see a clear  relationship. However,the majority points are located in the low plastic waster per capita and low mismanaged plastic waster per capita, suggesting that most continentes have small plastic waste and low mismanaged plastic waster.  
4.2 It looks like Asia and Africa have more points with higer mismanaged plastic waster per capita in the same levles of low plastic waster per capita, indicating these two countires have a stronger assocation between plastic waste and mismanaged plastic waster. It provides a more clear distinction for Asia and Africa. 
4.3 The plastic waste and costal populaton appear to be more strongly linearly associated because the points of the total  population are evenly distributed. 

```{r plastic-waste-mismanaged}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,   
    y = mismanaged_plastic_waste_per_cap,
  )
) +
  geom_point()
```

```{r plastic-waste-mismanaged-continent}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,   
    y = mismanaged_plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_point()
```

```{r plastic-waste-population-total}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,   
    y = total_pop,
    color = continent,
    fill = continent
  )
) +
  geom_point()
```

```{r plastic-waste-population-coastal}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,   
    y = coastal_pop,
    color = continent,
    fill = continent
  )
) +
  geom_point()
```

### Exercise 5

Most countries with lower coastal populations have higher plastic waste per capita. The black line does not show a linear relationship between plastic waste and coastal population proportion. The plastic waste starts to decrease at low coastal population proportions. Also, the shaded region only has a few points, indicating the uncertainty about the relationship in continents with higher coastal populations since it represents the confidence interval. Asia has the most uncertainly in the relationship between plastic waste and coastal population proportions, whereas most other countries are concentrated at low plastic waste and coastal population proportions. 

```{r recreate-viz}
filtered_data <- plastic_waste %>%
  filter(plastic_waste_per_cap <= 3)

ggplot(
  data = filtered_data,
  mapping = aes(
    y = plastic_waste_per_cap,   
    x = coastal_pop,
    color = continent,
  )
) +
  geom_point(alpha = 0.7) +
  geom_smooth(color = "black", method = 'loess', se= TRUE,) +  
  labs(
    title = "Plastic waste vs. coastal population proportion",
    subtitle = "By continent",
    x = "Coastal population proportion (Coastal / total population)",
    y = "Plastic waste per capita",
    color = "Continent")

```


